---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# pinr
[![Travis build status](https://travis-ci.org/mikmart/pinr.svg?branch=master)](https://travis-ci.org/mikmart/pinr)
[![Coverage status](https://codecov.io/gh/mikmart/pinr/branch/master/graph/badge.svg)](https://codecov.io/github/mikmart/pinr?branch=master)

The goal of **pinr** is to simplify working with data containing Finnish
personal identity codes (PINs). You can:

  * Check the validity of or extract information from PINs.
  * Use a heuristic to identify data columns that potentially contain PINs.
  * `pseudonymize()` existing data with a key file without manual joins.

## Installation

Currently you can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mikmart/pinr")
```

## Usage

The primary (pipe-friendly) utility function automates pseudonymizing columns containing PINs in your data:

```{r}
library(pinr)

df <- data.frame(pin = c("311280-888Y", "311280-888Y", "131052-308T"))
key <- data.frame(pin = c("311280-888Y", "131052-308T"), pid = c(1, 2))

pseudonymize(df, key, pid = pin)
```

The result is equivalent to looking up the `pid` from the key data frame, but **pinr** also takes care of managing the columns and naming in your data.

```{r}
key$pid[match(df$pin, key$pin)]
```

Rather than manually specifying columns containing PINs, you can also use a heuristic implemented in the `is_probably_pin()` function to `guess` which columns need to be pseudonymized:

```{r}
pseudonymize(df, key, guess = TRUE, remove = FALSE)
```

**pinr** also includes helpers for extracting data contained in the Finnish PINs, such as the date of birth and sex:

```{r}
pins <- c("311280-888Y", "131052-308T")

pin_dob(pins)

pin_sex(pins)
```


